---
title: "Inferring Vegetation Memory from Remote Sensing Data using novel Climate Reconstruction Products"
subtitle: "M.Sc. Thesis Defense"
author: "Erik Kusch"
date: "05/07/2019"
fontsize: 10pt
output: 
  beamer_presentation: 
    keep_tex: true
    toc: false # this is added through a later command
    slide_level: 3 # at how many pound signs (#) to assume slide title level
    includes:
      in_header: Style.tex
classoption: t
---

```{r setup, include=FALSE}
Dir.Pres <- getwd()
Dir.Base <- paste(c(strsplit(getwd(), "/")[[1]][1:(length(strsplit(getwd(), "/")[[1]])-2)]), collapse="/")

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, cache.lazy=TRUE, tidy.opts=list(width.cutoff=60),tidy=TRUE)
options(digits=2)

install.load.package <- function(x) {
  if (!require(x, character.only = TRUE))
    install.packages(x, repos='http://cran.us.r-project.org')
  require(x, character.only = TRUE)
}
package_vec <- c("knitr", "raster", "pracma", "gameofthrones", "ggplot2", "cowplot", 
                 "grid", "gridExtra", "xlsx", "kableExtra")
sapply(package_vec, install.load.package)
oldSource <- knit_hooks$get("source")
knit_hooks$set(source = function(x, options) {
  x <- oldSource(x, options)
  x <- ifelse(!is.null(options$ref), paste0("\\label{", options$ref,"}", x), x)
  ifelse(!is.null(options$codecap), paste0("\\captionof{chunk}{", options$codecap,"}", x), x)
})
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, cache.lazy=FALSE, tidy.opts=list(width.cutoff=80),tidy=TRUE, fig.height=5.5)
options(digits=4, width = 90)

setwd(Dir.Base)
source("Y - Codes/S0b_Directories.R") # setting directories
source("Y - Codes/S0c_Functions.R") # Loading miscellaneous functions
setwd(Dir.Pres)

SR_Titles <- list("NDVI", "Air Temperature", "Soil Moisture (0-7cm)", "Soil Moisture (7-28cm)", "Soil Moisture (28-100cm)", "Soil Moisture (100-255cm)")
setwd(Dir.KrigCov)
Elevation <- raster(list.files()[2], varname = "Elevation")
ElevationF <- raster(list.files()[1], varname = "Elevation")

col.signeg <- got(n = 100, alpha = 1, begin = 0, end = 1, direction = -1, option = "targaryen2")
col.sigpos <- got(n = 100, alpha = 1, begin = 0, end = 1, direction = -1, option = "tyrell")
col.nonsig <- colorRampPalette(c("grey"))(1)
col.lags <- got(n = 12, alpha = 1, begin = 0, end = 1, direction = 1, option = "daenerys")
```

```{r ModelRes, echo = FALSE}
ModelRes <- function(Region, SoilLayer, Plot = "ALL"){
  if(Plot == "ALL"){
    par(mfrow=c(2,2))
    smallplotxpos <- c(0.5,.93,.12,.17) # where to put colour scales
    smallplotxneg <- c(0.085,.5,.12,.17) # where to put colour scales
    smallplotsing <- c(.085, .93, .12, .17)
  }else{
    par(mfrow=c(1,1))
    smallplotxpos <- c(.5,.938,.03,.08) # where to put colour scales
    smallplotxneg <- c(.045,.5,.03,.08) # where to put colour scales
    smallplotsing <- c(.045, .938, .03, .08)
  }
  SR_Titles <- list("NDVI", "Air Temperature", "Soil Moisture (0-7cm)", "Soil Moisture (7-28cm)", "Soil Moisture (28-100cm)", "Soil Moisture (100-255cm)")
  Titles <- c('Soil Memory Lag', 'NDVI [t-1]', SR_Titles[SoilLayer+2], 'Air Temperature')
  Files <- list.files(Dir.Memory)[grep(list.files(Dir.Memory), pattern = Region)]
  Files <- Files[grep(Files, pattern = ".nc")]
  plot_ras <- brick(paste(Dir.Memory, Files[SoilLayer], sep="/"))
  if(Plot == "ALL" | Plot == 1){
  plot(plot_ras[[1]], col = col.lags, colNA = "black", legend=FALSE, axes=FALSE, main = Titles[1], cex.main = 3.5)
  plot(plot_ras[[1]], legend.only=TRUE, col=col.lags, smallplot= smallplotsing, horizontal = TRUE, axis.args=list(cex.axis=2.5))
}
  if(SoilLayer == 1){
    RasLay <- 4:6
  }else{
    RasLay <- c(5)
  }
  if(Plot != "ALL"){
    RasLay <- c(Plot+2)
  }
  if(Plot == "ALL" | Plot != 1){
   for(Plot in RasLay){
    Neg_ras <- plot_ras[[Plot]]
    Neg_ras[which(values(Neg_ras) >= 0)] <- NA
    Pos_ras <- plot_ras[[Plot]]
    Pos_ras[which(values(Pos_ras) < 0)] <- NA
    plot(Neg_ras, col=col.signeg, colNA = "black", legend=FALSE, axes=FALSE, main = Titles[(Plot-2)], cex.main = 3.5)
    if(Plot == 4 & Region == "Iberian Region"){
      plot(Pos_ras, col=col.sigpos, colNA = "black", legend=FALSE, axes=FALSE, add=TRUE)
    }else{
      plot(Pos_ras, col=col.sigpos, legend=FALSE, axes=FALSE, add=TRUE) 
    }
    plot(Neg_ras, legend.only=TRUE, col=col.signeg, colNA = "black", smallplot=smallplotxneg, horizontal = TRUE, axis.args=list(cex.axis=2.5))
    if(Plot == 4 & Region == "Iberian Region"){
      plot(Pos_ras, legend.only=TRUE, col=col.sigpos, smallplot=smallplotsing, horizontal = TRUE, axis.args=list(cex.axis=2.5))
    }else{
      plot(Pos_ras, legend.only=TRUE, col=col.sigpos, smallplot=smallplotxpos, horizontal = TRUE, axis.args=list(cex.axis=2.5))
    }
  } 
  }
}
```

```{r VarParRes, echo = FALSE}
VarParRes <- function(Region, SoilLayer, Plot = 1, Legend = TRUE, lim = FALSE){
  col.varpar1 <- got(n = 100, alpha = 1, begin = 0, end = 1, direction = -1, option = "wildfire")
col.varpar2 <- got(n = 100, alpha = 1, begin = 0, end = 1, direction = -1, option = "targaryen")
col.varpar3 <- got(n = 100, alpha = 1, begin = 0, end = 1, direction = -1, option = "jon_snow")
col.list <- list(col.varpar1, col.varpar2, col.varpar3) 
  Files <- list.files(Dir.Memory)[grep(list.files(Dir.Memory), pattern = Region)]
  Files <- Files[grep(Files, pattern = ".nc")]
  Alter_ras <- brick(paste(Dir.Memory, Files[SoilLayer], sep="/"))[[7:10]]
  Alter_ras[2] <- 0
  values(Alter_ras)[which(values(Alter_ras) < 0)] <- 0
  cells <- order(values(Alter_ras[[1]]))
  `%nin%` = Negate(`%in%`) # create a 'not in' statement
  cells <- cells[which(cells %nin% which(values(Alter_ras[[1]])>quantile(values(Alter_ras[[1]]), .95, na.rm = TRUE)))]
  plot_df <- data.frame(Data =NA, 
                        Cell =NA,
                        Variance =NA)
  Idents <- c("Total", "t-1", "Shared", "Qsoil")
  for(i in 1:4){
    if(i > 1){
      plot_df1 <- data.frame(Data = values(Alter_ras[[i]])[cells],
                             Cell = 1:length(cells), 
                             Variance = rep(Idents[i], length(cells)))
      plot_df <- rbind(plot_df, plot_df1)
    }
  }
  if(lim == TRUE){
    Lims <- c(0, 0.7)
  }else{
    Lims <- c(0, max(plot_df$Data, na.rm = TRUE))
  }
  plot_df <- na.omit(plot_df)
  if(Legend == TRUE){
   p <- ggplot(data = plot_df, aes(y = Data, x = Cell, fill = Variance)) + geom_bar(stat = "identity") + theme_bw(base_size= 45) + xlab("Raster Cells") + ylab("Variance") + scale_fill_manual(values=c(col.list[[3]][1], col.list[[2]][50], col.list[[1]][30])) + ylim(Lims)
  }else{
   p <- ggplot(data = plot_df, aes(y = Data, x = Cell, fill = Variance)) + geom_bar(stat = "identity") + theme_bw(base_size= 45) + xlab("Raster Cells") + ylab("Variance") + scale_fill_manual(values=c(col.list[[3]][1], col.list[[2]][50], col.list[[1]][30])) + theme(legend.position = "none") + ylim(Lims)
  }
  if(Plot == 1){
    return(p)
  }else{
   for(i in 2:4){
    col.varpar <- col.list[[i-1]]
    plot(Alter_ras[[i]], col=col.varpar, colNA = "black", legend=FALSE, axes=FALSE)
    plot(Alter_ras[[i]], legend.only=TRUE, col=col.varpar, smallplot=c(.05, .93, .03, .065), horizontal = TRUE, axis.args=list(cex.axis=2.5))
  } 
  }
}
```

```{r PatCausRes, echo = FALSE}
PatcausRes <- function(SoilLayer = 1, X, Y){
  Regions <- c("Iberian Region"
               , "Caatinga", "Australia"
               )
  Idents <- c("NDVI", "ENDVI", "Qsoil", "EQsoil", "Tair", "ETair", "Memory")
  Labs <- c("NDVI [t-1] Coefficients", "Mean NDVI ", paste("Qsoil", SoilLayer, " Coefficients", sep=""), paste("Mean  Qsoil", SoilLayer, sep=""), "Tair Coefficients", "Mean Tair", paste("Qsoil", SoilLayer, " Memory Lags", sep=""))
  # data
  plot_df <- data.frame(X = NA, Y = NA, Region = NA)
  for(i in 1:length(Regions)){
    Region <- Regions[i]
    Dir.Reg <- paste(Dir.Memory, "/", Region, "-1981_2015", sep="")
    Files <- list.files(Dir.Reg)[grep(list.files(Dir.Reg), pattern = ".nc")]
    Alter_ras <- brick(paste(Dir.Reg, Files[SoilLayer], sep="/"))
    Files <- list.files(Dir.ERA.Monthly)[grep(list.files(Dir.ERA.Monthly),
                                              pattern = Region)]
    Files <- Files[grep(Files, pattern = ".nc")]
    EQsoil <- Files[grep(Files, pattern = "Qsoil")]
    EQsoil <- brick(paste(Dir.ERA.Monthly, EQsoil[SoilLayer], sep="/"))
    EQsoil <- values(mean(EQsoil))
    ETair <- Files[grep(Files, pattern = "Tair")]
    ETair <- brick(paste(Dir.ERA.Monthly, ETair, sep="/"))
    ETair <- values(mean(ETair))
    ENDVI <- list.files(Dir.Gimms.Monthly)[grep(list.files(Dir.Gimms.Monthly),
                                                pattern =Region)]
    ENDVI <- brick(paste(Dir.Gimms.Monthly, ENDVI, sep="/"))
    ENDVI <- values(mean(ENDVI))
    values <- list(values(Alter_ras[[2]]), ENDVI,
                   values(Alter_ras[[3]]), EQsoil,
                   values(Alter_ras[[4]]), ETair,
                   values(Alter_ras[[1]])
                   )
    plot_df1 <- data.frame(X = values[[which(Idents == X)]],
                          Y = values[[which(Idents == Y)]],
                          Region = rep(Region, length(values[[which(Idents == X)]])))
    plot_df1 <- na.omit(plot_df1)
    plot_df <- rbind(plot_df, plot_df1)
  QsoilTitle <- c("(0-7cm)", "(7-28cm)", "(28-100cm)", "(100-255cm)")
  }
  plot_df <- na.omit(plot_df)
  # plotting
  p <- ggplot(plot_df, aes(x = X, y = Y, col = Region)) + 
    geom_point(shape = ".", alpha = 0.4, size = 0.5) + 
  theme_bw(base_size = 35) + ylab(Labs[which(Idents == Y)]) + xlab(Labs[which(Idents == X)]) +
  stat_smooth(method = "lm",level = 0.95, size = 3)+
  geom_hline(yintercept = 0, linetype="dotted") + 
    scale_color_manual(values=got(n = length(Regions), alpha = 1, direction = -1, option = "margaery"))
  if(X == "Memory"){
   p <- ggplot(plot_df, aes(x = as.factor(X), y = Y, col = Region)) + geom_boxplot() + theme_bw(base_size = 35) + ylab(Labs[which(Idents == Y)]) + xlab(Labs[which(Idents == X)]) + 
  geom_hline(yintercept = 0, linetype="dotted") + 
    scale_color_manual(values=got(n = length(Regions), alpha = 1, direction = -1, option = "margaery")) + 
     stat_smooth(method = "lm",level = 0.95, aes(group=plot_df$Region), size = 3)
  }
  if(Y == "Memory"){
   p <- ggplot(plot_df, aes(x = as.factor(Y), y = X, col = Region)) + geom_boxplot() + theme_bw(base_size = 35) + xlab(Labs[which(Idents == Y)]) + ylab(Labs[which(Idents == X)]) + coord_flip() +
  geom_hline(yintercept = 0, linetype="dotted") + 
    scale_color_manual(values=got(n = length(Regions), alpha = 1, direction = -1, option = "margaery")) + 
       stat_smooth(method = "lm",level = 0.95, aes(group=plot_df$Region), size = 3)
  }
  print(p)
  # Regressions
  Output <- as.list(rep(NA, 4*length(Regions)))
  pos <- 1
  for(i in 1:length(Regions)){
    Output[[pos]] <- summary(lm(Y ~ X, data = plot_df[which(plot_df$Region == Regions[i]),]))[["coefficients"]][1,1]
    pos <- pos+1
    Output[[pos]] <- summary(lm(Y ~ X, data = plot_df[which(plot_df$Region == Regions[i]),]))[["coefficients"]][1,4]
    pos <- pos+1
    Output[[pos]] <- summary(lm(Y ~ X, data = plot_df[which(plot_df$Region == Regions[i]),]))[["coefficients"]][2,1]
    pos <- pos+1
    Output[[pos]] <- summary(lm(Y ~ X, data = plot_df[which(plot_df$Region == Regions[i]),]))[["coefficients"]][2,4]
    pos <- pos+1
  }
  return(Output)
}
```

```{r COMRes, echo = FALSE}
Comres <- function(Variable, Region, Legend = FALSE){
  col.tair <- got(n = 1, alpha = 1, begin = 0, end = 1, direction = -1, option = "tully")
col.ndvi <- got(n = 1, alpha = 1, begin = 0, end = 1, direction = 1, option = "tyrell")
col.qsoil <- got(n = 1, alpha = 1, begin = 0, end = 1, direction = 1, option = "white_walkers")
col.mem <- got(n = 1, alpha = 1, begin = 0, end = 1, direction = -1, option = "greyjoy")
  VarHold <- Variable
  if(Variable == "FSC-1" | Variable == "FSC-2"){
    Variable <- "FastSlow"
  }
  ## Memory Data
  # Memory
  Dir.Reg <- paste(Dir.Memory, "/", Region, "-1981_2015", sep="")
  Files <- list.files(Dir.Reg)[grep(list.files(Dir.Reg), pattern = ".nc")]
  SoilLayer <- 1
  Memory_ras <- brick(paste(Dir.Reg, Files[SoilLayer], sep="/"))[[1]]
  # t-1
  Dir.Reg <- paste(Dir.Memory, "/", Region, "-1981_2015", sep="")
  Files <- list.files(Dir.Reg)[grep(list.files(Dir.Reg), pattern = ".nc")]
  SoilLayer <- 1
  NDVI_ras <- brick(paste(Dir.Reg, Files[SoilLayer], sep="/"))[[2]]
  # Qsoil
  Dir.Reg <- paste(Dir.Memory, "/", Region, "-1981_2015", sep="")
  Files <- list.files(Dir.Reg)[grep(list.files(Dir.Reg), pattern = ".nc")]
  SoilLayer <- 1
  Qsoil_ras <- brick(paste(Dir.Reg, Files[SoilLayer], sep="/"))[[3]]
  # Tair
  Dir.Reg <- paste(Dir.Memory, "/", Region, "-1981_2015", sep="")
  Files <- list.files(Dir.Reg)[grep(list.files(Dir.Reg), pattern = ".nc")]
  SoilLayer <- 1
  Tair_ras <- brick(paste(Dir.Reg, Files[SoilLayer], sep="/"))[[4]]
  ## COMPADRE data
  Dir.Comp <- paste(Dir.Compadre, Variable, sep="/")
  Compad_ras <- list.files(Dir.Comp)[grep(list.files(Dir.Comp), pattern = Region)]
  if(VarHold != "FSC-2"){
    Compad_ras <- raster(paste(Dir.Comp, Compad_ras, sep="/"))[[1]]
  }else{
    Compad_ras <- brick(paste(Dir.Comp, Compad_ras, sep="/"))[[2]]
  }
  values(Compad_ras)[which(values(Compad_ras) > quantile(values(Compad_ras), .66, na.rm = TRUE))] <- quantile(values(Compad_ras), .66, na.rm = TRUE)
  if(VarHold == "FSC-1" | VarHold == "FSC-2"){
   values(Compad_ras)[which(values(Compad_ras) < quantile(values(Compad_ras), .05, na.rm = TRUE))] <- quantile(values(Compad_ras), .05, na.rm = TRUE) 
  }
  plot_df <- data.frame(
    Data = c(values(Memory_ras), values(NDVI_ras), values(Qsoil_ras), values(Tair_ras)), 
    Identifiers = rep(c("Lag", "t-1", "Qsoil1", "Tair"), each = length(values(Memory_ras))),
   Compadre = rep(values(Compad_ras), 4)
  )
  Variable <- VarHold
  plot_df <- na.omit(plot_df)
  Output <- as.list(rep(NA,16))
  Idents <- c("t-1", "Tair", "Qsoil1", "Lag")
  for(i in 0:(length(Idents)-1)){
    Output[[(4*i+1)]] <- summary(lm(Data ~ Compadre, data = plot_df[which(plot_df$Identifiers == Idents[(i+1)]),]))[["coefficients"]][1,1]
    Output[[(4*i+2)]] <- summary(lm(Data ~ Compadre, data = plot_df[which(plot_df$Identifiers == Idents[(i+1)]),]))[["coefficients"]][1,4]
    Output[[(4*i+3)]] <- summary(lm(Data ~ Compadre, data = plot_df[which(plot_df$Identifiers == Idents[(i+1)]),]))[["coefficients"]][2,1]
    Output[[(4*i+4)]] <- summary(lm(Data ~ Compadre, data = plot_df[which(plot_df$Identifiers == Idents[(i+1)]),]))[["coefficients"]][2,4]
  }
  Linesa <- rep(1,4)
  Linesa[which(unlist(Output)[c(4,8,12,16)] > .05)] <- 2
  Lines <- c(Linesa[4], Linesa[1], Linesa[3], Linesa[2])
  Lines <- rep(Lines, each = length(which(plot_df$Identifiers == "Lag")))
  Lines <- as.factor(Lines)
  plot_df <- cbind(plot_df, Lines)
  if(length(levels(Lines)) == 2){
    if(Legend == TRUE){
   plot <- ggplot(data = plot_df, aes(x = Compadre, y = Data, col = Identifiers, linetype = Lines)) + geom_point(alpha = 0.5, size = 3.5) + theme_bw(base_size = 35) +  xlab(paste("COMPADRE", Variable)) + ylab("Vegetation Response Coefficients") + geom_hline(yintercept = 0, linetype="dotted") + stat_smooth(method = "lm", level = 0.66, aes(linetype = Lines), size = 3) + scale_color_manual(values=c(col.mem, col.qsoil, col.ndvi, col.tair)) + 
     labs(linetype="p < .05", colour="Identifiers") + scale_linetype_manual(values = c(1, 2), labels = c("Yes", "No")) + guides(colour = guide_legend(ncol=2, override.aes = list(size = 5)),linetype = guide_legend(override.aes = list(size = .5)))
  }else{
   plot <- ggplot(data = plot_df, aes(x = Compadre, y = Data, col = Identifiers, linetype = Lines)) + geom_point(alpha = 0.5, size = 3.5) + theme_bw(base_size = 35) +  xlab(paste("COMPADRE", Variable)) + ylab("Vegetation Response Coefficients") + geom_hline(yintercept = 0, linetype="dotted") + stat_smooth(method = "lm", level = 0.66, aes(linetype = Lines), size = 3) + scale_color_manual(values=c(col.mem, col.qsoil, col.ndvi, col.tair)) + theme(legend.position = "none")
  }
  }else{
    if(Legend == TRUE){
   plot <- ggplot(data = plot_df, aes(x = Compadre, y = Data, col = Identifiers, linetype = Lines)) + geom_point(alpha = 0.5, size = 3.5) + theme_bw(base_size = 35) +  xlab(paste("COMPADRE", Variable)) + ylab("Vegetation Response Coefficients") + geom_hline(yintercept = 0, linetype="dotted") + stat_smooth(method = "lm", level = 0.66, linetype = 2, size = 3) + scale_color_manual(values=c(col.mem, col.qsoil, col.ndvi, col.tair)) + 
     labs(linetype="p < .05", colour="Identifiers") + scale_linetype_manual(values = c(1, 2), labels = c("Yes", "No")) + guides(colour = guide_legend(ncol=2, override.aes = list(size = 5)),linetype = guide_legend(override.aes = list(size = .5)))
  }else{
   plot <- ggplot(data = plot_df, aes(x = Compadre, y = Data, col = Identifiers, linetype = Lines)) + geom_point(alpha = 0.5, size = 3.5) + theme_bw(base_size = 35) +  xlab(paste("COMPADRE", Variable)) + ylab("Vegetation Response Coefficients") + geom_hline(yintercept = 0, linetype="dotted") + stat_smooth(method = "lm", level = 0.66, linetype = 2, size = 3) + scale_color_manual(values=c(col.mem, col.qsoil, col.ndvi, col.tair)) + theme(legend.position = "none")
  }
  }
  return(list(plot,Output,dim(plot_df)[1]/4))
}
```

\tableofcontents

# Background
## Motivation
### What is this?
\centering \textbf{Vegetation Memory} is the effect of \textbf{antecedent ecosystem/environmental anomalies} on \textbf{current vegetation performance}\cite{Liu2018}.

\vspace{.4cm}
\begincols[T]
  \begincol{.5\linewidth}
  \pause
  \textbf{Components of Memory:}\cite{Ogle2015}
    \begin{enumerate}
      \item \textit{Intrinsic Memory} (e.g. antecedent vegetation characteristics)\cite{DeKeersmaecker2015,Ogle2015}
      \item \textit{Extrinsic Memory} (antecedent climate characteristics)\cite{DeKeersmaecker2015,Vicente-Serrano2013,Ogle2015}
    \end{enumerate}
    \pause
    \vspace{0.3cm}
    \begin{tcolorbox}[colback=bondiblue!5,colframe=bondiblue!40!black,title= ]
    \centering Can we \textbf{distinguish} between \textbf{intrinsic} and \textbf{extrinsic} memory effects?
    \end{tcolorbox}
  \endcol
  \begincol{.5\linewidth}
  \pause
  \textbf{Explaining Memory:}
  \begin{enumerate}
      \item \textit{Causal pathways} remain poorly understood \cite{Smith2017}
      \item Expressions of \textit{Plant Function} as a possible solution
    \end{enumerate}
    \pause
    \vspace{0.3cm}
    \begin{tcolorbox}[colback=bondiblue!5,colframe=bondiblue!40!black,title= ]
    \centering What \textbf{biological traits cause} areas to exert        \textbf{intrinsic} and \textbf{extrinsic memory}?
    \end{tcolorbox}
  \endcol
\endcols

### Intrinsic vs. Extrinsic Memory Components
\includegraphics[width=1\linewidth]{Figures/Scheme_VegMem1.jpg}

\pause
\vspace{.5cm}
$\rightarrow$ Big emphasis on \textbf{dryland regions} due to demonstrated vegetation memory effects\cite{Seddon2016,Liu2018,DeKeersmaecker2015,Vicente-Serrano2013}, and the strong dependence of dryland vegetation on local water regimes\cite{Smith2017}

### Why do this?
\centering \textbf{Intrinsic vegetation Memory} proposed to be a \textit{proxy} of \textbf{engineering resilience in ecosystems} (\textit{high memory} $\sim$ \textit{low resilience})\cite{DeKeersmaecker2015}.

\vspace{.2cm}
  \includegraphics[width=1\linewidth]{Figures/DeKeers.png}
  \vspace{-.4cm}
  \pause
      \begin{tcolorbox}[colback=bondiblue!5,colframe=bondiblue!40!black,title= ]
    \centering How \textbf{valid} is this \textbf{assumption?}
    \end{tcolorbox}

## Dryland Vegetation Memory
### Study Regions
\vspace{-.2cm}

\textbf{The Iberian Region}
\vspace{.1cm}
\centering \includegraphics[height=1\textheight]{Figures/Ombro_EU_Detail.png}
\pause
\vspace{.3cm}
\raggedright Additional Study Regions:
\vspace{-.73cm}
\begincols
  \begincol{.33\linewidth}
  \begin{itemize}
\item Caatinga, Brazil
\end{itemize}
  \endcol
  \begincol{.33\linewidth}
  \begin{itemize}
\item Australia
\end{itemize}
  \endcol
  \begincol{.33\linewidth}
  \begin{itemize}
\item Contiguous US
\end{itemize}
  \endcol
\endcols


<!-- \only<1>{\centering \vspace{0.2cm} -->
<!-- \textbf{The Iberian Region} \vspace{0.2cm} -->
<!-- \includegraphics[height=1\textheight]{Figures/Ombro_EU_Detail.png}} -->
<!-- \only<2>{\centering \includegraphics[height=.85\textheight]{Figures/Ombro_SA.png}} -->
<!-- \only<3>{\centering \includegraphics[height=.85\textheight]{Figures/Ombro_AU.png}} -->

<!-- ### Study Outline -->
<!-- \vspace{.2cm} -->
<!-- \includegraphics[width=1\linewidth]{Figures/ThesisOutlinebw.png} -->

# Allocating and Preparing Data
## Vegetation Data
### Normalised Difference Vegetation Index (NDVI)
<!-- \centering We are using the \textbf{GIMMS NDVI3g} (v.1).   -->

<!-- \begin{equation} -->
<!-- NDVI = \frac{\rho\textsubscript{NIR} - \rho\textsubscript{RED}}{\rho\textsubscript{NIR} + \rho\textsubscript{RED}} -->
<!-- \end{equation} -->

<!-- \raggedright -->
\begin{itemize}
  \item \textit{Biological Relevance:} Proxy of biomass and vegetation cover
  \item \textit{Comparability:} Has been used in other studies of vegetation memory
\end{itemize}

```{r NDVIOverview, echo = FALSE, fig.width = 18, fig.height = 9.9, cache.lazy=FALSE}
setwd(Dir.Gimms.Monthly)
NDVIMean_ras <- mean(brick(list.files()[1:7]), na.rm = TRUE)
plot(NDVIMean_ras, colNA = "black", main = "Mean NDVI 1982 - 2015", cex.main = 2, 
     legend.width = 1.5, legend.shrink=1, axis.args=list(cex.axis=1.5), cex.axis = 1.5
     )
```
<!-- \pause -->
<!-- \begin{itemize} -->
<!-- \item \textbf{Core Characteristics:} -->
<!--   \begin{itemize} -->
<!--   \item \textit{Resolution:} $0.083^\circ \sim 9.27km \times 9.27km$ -->
<!--   \item \textit{Availability:} 1982 - 2015 -->
<!--   \end{itemize} -->
<!-- \end{itemize} -->

## Climate Data
### ERA5 & Climate Variables
\begin{itemize}
\item \textbf{Why:}
  \begin{itemize}
  \item Applicable globally
  \item Gap-less time series
  \item More sophisticated approach than previously utilised:
    \begin{itemize}
    \item Worldclim - Superior Temporal Resolution (superior resolving of climate extremes)
    \item CRU - Superior Spatial Resolution
    \end{itemize}
  \item \textit{Climatological Relevance:} Assessment of climate uncertainty possible
  \end{itemize}
\vspace{.5cm}
\visible<2->{
\item \textbf{Soil Moisture} - $Qsoil$
}
  \begin{itemize}
  \visible<3->{
  \item \textit{Why:} Soil moisture effects on vegetation may indicate if/how the local ecosystem may deal with drought stress\cite{Rudgers2018,Smith2017,Papagiannopoulou2017a}.
  \item \textit{How:} as different layers of depth ($Qsoil1$ - $Qsoil4$)
  }
  \end{itemize}
\vspace{.2cm}
\visible<2->{
\item \textbf{Air Temperature} - $Tair$
}
  \begin{itemize}
  \visible<4->{
  \item \textit{Why:} Temperature drives plant physiology and drives levels of aridity\cite{Rudgers2018,DeKeersmaecker2015}.
  \item \textit{How:} As one single layer (2m above ground)
  }
  \end{itemize}
\end{itemize}

### Kriging
\includegraphics[width=1\linewidth]{Figures/Kriging.jpg}

## Plant Functional Data
### COMPADRE Data
\textbf{Life History Traits} (LHTs)
  \begin{itemize}
  \pause
  \item \textbf{Why:}
    \begin{itemize}
    \item \textit{Biological Relevance:} Indices of plant behaviour through time
    \item \textit{Comparability:} Capture much of natural life strategy variation\cite{Salguero-Gomez2015}
     \end{itemize}
   \pause
   \vspace{.2cm}
   \item \textbf{Core Measures:}
     \begin{itemize}
     \item \textit{Fast-Slow Continuum (FSC):} Capture over 60\% of the variation in plant life history strategies
     \begin{itemize}
       \item FSC-1: Life History Speed
       \item FSC-2: Reproductive Strategy/Output
       \end{itemize}
     \item \textit{Reactivity:} Instantaneous biological responses
     \end{itemize}
  \end{itemize}
\vspace{1cm}
\pause
Expressions of \textbf{Plant Functional Traits} (PFTs) were tested against vegetation memory but require further research aso of this point in time.
   
<!-- \textbf{Plant Functional Traits} (PFTs) -->
<!--     \begin{itemize} -->
<!--     \pause -->
<!--     \item \textbf{Why:} -->
<!--       \begin{itemize} -->
<!--       \item \textit{Biological Relevance:} Snapshots of functional reality -->
<!--       \item \textit{Comparability:} Have widely been used in functional ecology\cite{Ordonez2015,Diaz2015} -->
<!--       \end{itemize} -->
<!--     \pause -->
<!--     \vspace{.2cm} -->
<!--     \item \textbf{Core Measures:} -->
<!--       \begin{itemize} -->
<!--       \item \textit{Vegetative Height:} Ability to pre-empt light -->
<!--       \item \textit{Leaf Nitrogen Content:} Measure of photosynthetic potential and nitrogen acquisition -->
<!--       \end{itemize} -->
<!--     \end{itemize} -->

# Delineating Vegetation Memory
### Pixyel-Wise Model Building
\begin{itemize}
\item Linear detrending  
\item Z-Scores:
\end{itemize}
\small
\begin{equation}
Anomaly_{i} = \frac{Detrended_{i} - \overline{Detrended_{month}}}{SD_{Detrended, month}}
\end{equation}
\normalsize
\begin{itemize}
\item Calculate:
  \begin{itemize}
  \item $t-1$ lag for NDVI
  \item Cummulative lags for $Qsoil$ data
  \end{itemize}
\item Set NDVI anomalies to `NA` in months for which $Thresholds_{i} < 0.1$ with $Thresholds_{i} = \overline{Raw_{NDVI, month}}$
\item PCA regression and model selection:
\end{itemize}
\small
\begin{equation}
NDVI_t = \beta_{t-1}*NDVI_{[t-1]} + \beta_{Qsoil}*Qsoil_{k;m} + \beta_{Tair}*Tair_{t}
\end{equation}
<!-- \vspace{.2cm} -->
<!-- \centering \includegraphics[width=1\linewidth]{Figures/ModScheme.jpg} -->

<!-- ### Examplary Data -->
<!--  \vspace{.4cm} -->
<!-- \centering \includegraphics[width=.9\linewidth]{Figures/ExMod.jpg} -->

<!-- ### PCA Regression -->
<!-- \centering \includegraphics[width=1\linewidth]{Figures/PCAREG.jpg} -->
<!--  \pause -->
<!-- \raggedright \textbf{Model Comparisons:} -->
<!-- \begin{itemize} -->
<!-- \item \textit{Which variable exerts the greatest influence on vegetation anomalies?}  -->
<!-- \item \textit{Which soil moisture layer is the most biologically influential?}  -->
<!-- \end{itemize} -->

<!-- ## Causes of Vegetation Memory -->
<!-- ### Vegetation Memory Adaptation -->
<!-- ### Plant Function -->


# Results
## Coefficients of Vegetation Memory

### Vegetation Memory Coefficients
```{r IBR, echo = FALSE, fig.width = 18, fig.height = 9.9, cache = TRUE, cache.lazy=FALSE}
ModelRes(Region = "Iberian Region", SoilLayer = 1)
```
\pause
\centering $\rightarrow$ Is \textbf{intrinsic} memory really \textit{intrinsic}?

### Distinguishing Intrinsic and Extrinsic Memory
\centering \textbf{Qsoil1} is the \textbf{most informative} of the \textbf{soil moisture} layers!
```{r IBR3, echo = FALSE, fig.width = 18, fig.height = 10}
VarParRes(Region = "Iberian Region", SoilLayer = 1, Plot = 1)
```

### Identifying Underlying Extrinsic Patterns I
\centering Uniform **$NDVI[t-1]$ effect** across Australia **contrasts** with **other studies**.
\begincols
  \begincol{.5\linewidth}
  ```{r AU1, echo = FALSE, fig.width = 18, fig.height = 20, cache =     TRUE, cache.lazy=FALSE}
  ModelRes(Region = "Australia", SoilLayer = 1, Plot = 2)
  ```
  \endcol
  \begincol{.5\linewidth}
  \vspace{-6.75cm}
  
  Intrinsic Memory by Seddon et al.\cite{Seddon2016}:  
  \vspace{.1cm}
  \centering \includegraphics[width=1\linewidth]{Figures/SeddonAU.jpg}
  
  \endcol
\endcols

### Identifying Underlying Extrinsic Patterns II
\centering Uniform **$NDVI[t-1]$ effect** across Australia **contrasts** with **other studies**.

\vspace{-.41cm}

\begincols
  \begincol{.5\linewidth}
  ```{r AU3, echo = FALSE, fig.width = 18, fig.height = 20, cache =     TRUE, cache.lazy=FALSE}
  ModelRes(Region = "Australia", SoilLayer = 1, Plot = 1)
  ```
  
  \endcol
  \begincol{.5\linewidth}
  
  Intrinsic Memory by Seddon et al.\cite{Seddon2016}:  
  \vspace{.1cm}
  \centering \includegraphics[width=1\linewidth]{Figures/SeddonAU.jpg}   
  \vspace{-.1cm}
  
  \only<2>{
    \begin{tcolorbox}[colback=green!5,colframe=green!40!black,title= ]
    \centering \textbf{Previous $t-1$ variation } can be \textbf{understood through extrinsic vegetation memory}.
    \end{tcolorbox}
  }
  \endcol
\endcols

## Regional Aspects of Vegetation Memory
### Vegetation Memory Adaptation
```{r VegMemSens5, echo = FALSE, fig.width = 18, fig.height = 8.5}
PatCauseMem <- PatcausRes(SoilLayer = 1, X = "Memory", Y = "NDVI")
```

\pause

\begin{tcolorbox}[colback=green!5,colframe=green!40!black,title= ]
\centering Relationship of $t-1$ coefficient and extrinsic vegetation memory length is not uniform within or between study regions.
\end{tcolorbox}


## Functional Aspects to Vegetation Memory
### Plant Function I
\centering Linking **plant functional traits** and vegetation memory proved **non-conclusive** but **life history traits** showed **interesting patterns**:
```{r IBCOMPAD1, echo = FALSE, fig.width = 17, fig.height = 8.5, cache = TRUE, cache.lazy=FALSE, out.extra=''}
leg2 <- Comres(Region = "Iberian Region", Variable = "FSC-2", Legend = TRUE)
leg2[[1]]
```

### Plant Function II
```{r IBCOMPAD2, echo = FALSE, fig.width = 17, fig.height = 8.5, cache = TRUE, cache.lazy=FALSE, out.extra=''}
leg5 <- Comres(Region = "Iberian Region", Variable = "Reactivity", Legend = TRUE)
leg5[[1]]
```

\pause

\begin{tcolorbox}[colback=green!5,colframe=green!40!black,title= ]
\centering Some LHTs can be linked to some vegetation memory characteristics.
\end{tcolorbox}

# Conclusion
### Conclusion  
\begincols[T]
  \begincol{.5\linewidth}
  \textbf{Summary:}
  \begin{tcolorbox}[colback=violet!5,colframe=violet!40!black,title= ]
  \centering Extrinsic memory should not be neglected in favour of intrinsic memory. 
  \end{tcolorbox}
  \endcol
  \begincol{.5\linewidth}
  ```{r conclus1, echo = FALSE, fig.width = 18, fig.height = 20, cache = TRUE, cache.lazy=FALSE}
  ModelRes(Region = "Iberian Region", SoilLayer = 1, Plot = 2)
  ```
  \endcol
\endcols

### Conclusion  
\begincols[T]
  \begincol{.5\linewidth}
  \textbf{Summary:}
  \begin{tcolorbox}[colback=violet!5,colframe=violet!40!black,title= ]
  \centering Extrinsic memory should not be neglected in favour of intrinsic memory. 
  \end{tcolorbox}
  \begin{enumerate}
  \vspace{.2cm}
  \item Intrinsic vegetation memory as a proxy for engineering resilience may be an oversimplification.
  \end{enumerate}
  \endcol
  \begincol{.5\linewidth}
  ```{r conclus2, echo = FALSE, fig.width = 18, fig.height = 20, cache = TRUE, cache.lazy=FALSE}
  ModelRes(Region = "Iberian Region", SoilLayer = 1, Plot = 3)
  ```
  \endcol
\endcols

### Conclusion  
\begincols[T]
  \begincol{.5\linewidth}
  \textbf{Summary:}
  \only<1-2>{
  \begin{tcolorbox}[colback=violet!5,colframe=violet!40!black,title= ]
  \centering Extrinsic memory should not be neglected in favour of intrinsic memory. 
  \end{tcolorbox}
  }
  \visible<1-3>{
  \begin{enumerate}
  \vspace{.2cm}
  \item Intrinsic vegetation memory as a proxy for engineering resilience may be an oversimplification.
  \end{enumerate}
  }
  \endcol
  \begincol{.5\linewidth}
  \raggedleft
  ```{r conclus3, echo = FALSE, fig.width = 18, fig.height = 16}
  VarParRes(Region = "Iberian Region", SoilLayer = 1, Plot = 1)
  ```
  \visible<2->{
  Distinguishing intrinsic and extrinsic memory components remains challenging.
  }
  \endcol
\endcols

### Conclusion  
\begincols[T]
  \begincol{.5\linewidth}
  \textbf{Summary:}
  \begin{enumerate}
  \vspace{.2cm}
  \item Intrinsic vegetation memory as a proxy for engineering resilience may be an oversimplification.
  \end{enumerate}
  \vspace{.6cm}
  \begin{tcolorbox}[colback=violet!5,colframe=violet!40!black,title= ]
  \centering Interactions of vegetation memory characteristics are region-dependant.
  \end{tcolorbox}
  \visible<2>{
  \begin{enumerate}
  \setcounter{enumi}{1}
  \vspace{.2cm}
  \item Vegetation memory processes differ greatly between regions.
  \end{enumerate}
  }
  \endcol
  \begincol{.5\linewidth}
  ```{r conclus4, echo = FALSE, fig.width = 18, fig.height = 20}
  PatCauseMem <- PatcausRes(SoilLayer = 1, X = "Memory", Y = "NDVI")
  ```
  \endcol
\endcols

### Conclusion  
\begincols[T]
  \begincol{.5\linewidth}
  \textbf{Summary:}
  \begin{enumerate}
  \vspace{.2cm}
  \item Intrinsic vegetation memory as a proxy for engineering resilience may be an oversimplification.
  \vspace{.2cm}
  \item Vegetation memory processes differ greatly between regions.
  \end{enumerate}
  \only<1-2>{
  \vspace{.6cm}
  \begin{tcolorbox}[colback=violet!5,colframe=violet!40!black,title= ]
  \centering Some measures of life history strategies are related to vegetation memory characteristics.
  \end{tcolorbox}
  }
  \visible<3>{
  \begin{enumerate}
  \setcounter{enumi}{2}
  \vspace{.2cm}
  \item The relationship of plant function and vegetation memory requires further research.
  \end{enumerate}
  }
  \endcol
  \begincol{.5\linewidth}
  \raggedleft
  ```{r conclus5, echo = FALSE, fig.width = 18, fig.height = 16}
  leg5[[1]]
  ```
  \visible<2->{
  Challenging to establish direct proxies of either intrinsic or extrinsic vegetation memory components.
  }
  \endcol
\endcols

### Conclusion  
\begincols[T]
  \begincol{.5\linewidth}
  \textbf{Summary:}
  \begin{enumerate}
  \vspace{.2cm}
  \item Intrinsic vegetation memory as a proxy for engineering resilience may be an oversimplification.
  \vspace{.2cm}
  \item Vegetation memory processes differ greatly between regions.
  \vspace{.2cm}
  \item The relationship of plant function and vegetation memory requires further research.
  \end{enumerate}
  \endcol
  \begincol{.5\linewidth}
  \textbf{Future Directions:}
  \begin{enumerate}
  \vspace{.2cm} \pause
  \item Identify additional relevant climate parameters (e.g. cloud cover and soil type)
  \vspace{.2cm} \pause
  \item Extend analyses beyond dryland regions
  \vspace{.2cm} \pause
  \item Analyse changes in vegetation memory over time
  \vspace{.2cm} \pause
  \item Use additional plant functional data (e.g. BIEN data base)
  \vspace{.2cm} \pause
  \item Implement advanced mapping methods for plant functional data
  \end{enumerate}
  \endcol
\endcols



### Sources
\tiny
\MakeBib